Elephant Seal Age JAE Intraspecific Variation

### 2024-06-07 paper/paperv3.qmd

### H1: Age distribution and breeding status

@@

breed_age_mod <- glmer(
  observed_int ~ poly(age10, 2, raw = TRUE) + (1 | animalID) + (1 | year_fct),
  sealdat,
  family = "binomial",
  control = glmerControl(optimizer = "bobyqa")
)

@@

```{r}
#| label: fig-breeding-mod

@@

# Create POPULATION-LEVEL (i.e. without rand effects) predictions
breed_pred_pop <- ggpredict(
  breed_age_mod, 
  terms = "age10 [all]",
  type = "fixed",
  interval = "confidence"
) %>% 
  as_tibble() %>% 
  mutate(age = x * 10) %>% 
  filter(age <= 20)

# Create WEIGHTED-AVERAGE predictions using random effects
seal_years <- count(sealdat, year_fct)
breed_pred_wgt <- ggpredict(
  breed_age_mod,
  terms = c("age10 [all]", "year_fct [all]"),
  type = "random",
  interval = "confidence"
) %>% 
  as_tibble() %>% 
  mutate(age = x * 10,
         year_fct = group) %>% 
  filter(age <= 20) %>% 
  left_join(seal_years, by = "year_fct") %>% 
  group_by(age) %>% 
  summarize(predicted = weighted.mean(predicted, n))

ggplot(breed_pred_pop, aes(age, predicted)) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), 
              alpha = 0.2) +
  geom_line(linewidth = 1.2, color = "grey50") +
  geom_line(data = breed_pred_wgt, 
            color = "cornflowerblue", 
            linewidth = 1.2) +
  geom_pointrange(aes(y = perc_breed, 
                      ymin = perc_breed - se, 
                      ymax = perc_breed + se),
                  breed_nonbreed) +
  scale_y_continuous("Breeding", labels = scales::percent) +
  labs(x = "Age (years)") +
  theme(legend.position = "none")

@@

##NOTE: Need to take out the instances where recruitment > survival

# Isolate pup survival/recruitment 
surv_recr_data <- sealdat %>% 
  drop_na(pupID) %>% 
  # Only pups born 2016 or earlier (otherwise survival biased)
  filter(pupyearborn <= 2016) %>% 
  mutate(pup_surv_int = as.integer(pup_survived),
         pup_recr_int = as.integer(puprecruited)) %>% 
  # Only keep maternal age classes with 5+ births, which drops moms 19+
  group_by(age) %>% 
  mutate(n_pups = n()) %>% 
  ungroup() %>%
  filter(n_pups >= 5)

# Fit pup *survival* model
pup_surv_mod <- glmer(
  pup_surv_int ~ poly(age10, 2, raw = TRUE) + (1 | animalID) + (1 | year_fct),
  surv_recr_data,
  family = "binomial",
  control = glmerControl(optimizer = "bobyqa")
)

# Fit pup *recruitment* model (female offspring only)
pup_recr_mod <- glmer(
  pup_recr_int ~ poly(age10, 2, raw = TRUE) + (1 | animalID) + (1 | year_fct),
  filter(surv_recr_data, pupsex == "F"),
  family = "binomial",
  control = glmerControl(optimizer = "bobyqa")
)

@@

# Create pup survival POPULATION-LEVEL (i.e. without rand effects) predictions
surv_pred_pop <- ggpredict(
  pup_surv_mod, 
  terms = "age10 [all]",
  type = "fixed",
  interval = "confidence"
) %>% 
  as_tibble() %>% 
  mutate(age = x * 10) %>% 
  filter(age <= 20)

# Create pup survival WEIGHTED-AVERAGE predictions using random effects
seal_years <- count(surv_recr_data, year_fct)
surv_pred_wgt <- ggpredict(
  pup_surv_mod,
  terms = c("age10 [all]", "year_fct [all]"),
  type = "random",
  interval = "confidence"
) %>% 
  as_tibble() %>% 
  mutate(age = x * 10,
         year_fct = group) %>% 
  filter(age <= 20) %>% 
  left_join(seal_years, by = "year_fct") %>% 
  group_by(age) %>% 
  summarize(predicted = weighted.mean(predicted, n))

fig_surv <- ggplot(surv_pred_pop, aes(age, predicted)) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), 
              alpha = 0.2) +
  geom_line(linewidth = 1.2, color = "grey50") +
  geom_line(data = surv_pred_wgt, 
            color = "cornflowerblue", 
            linewidth = 1.2) +
  geom_pointrange(aes(y = perc_surv, 
                      ymin = perc_surv - se_surv, 
                      ymax = perc_surv + se_surv),
                  surv_recr_summ) +
  scale_y_continuous("Offspring survival", 
                     labels = scales::percent,
                     limits = c(0, 0.62)) +
  labs(x = "Age (years)") +
  theme(legend.position = "none")

# Create pup recruit POPULATION-LEVEL (i.e. without rand effects) predictions
recr_pred_pop <- ggpredict(
  pup_recr_mod, 
  terms = "age10 [all]",
  type = "fixed",
  interval = "confidence"
) %>% 
  as_tibble() %>% 
  mutate(age = x * 10) %>% 
  filter(age <= 20)

# Create pup recruit WEIGHTED-AVERAGE predictions using random effects
seal_years <- count(filter(surv_recr_data, pupsex == "F"), year_fct)
recr_pred_wgt <- ggpredict(
  pup_recr_mod,
  terms = c("age10 [all]", "year_fct [all]"),
  type = "random",
  interval = "confidence"
) %>% 
  as_tibble() %>% 
  mutate(age = x * 10,
         year_fct = group) %>% 
  filter(age <= 20) %>% 
  left_join(seal_years, by = "year_fct") %>% 
  group_by(age) %>% 
  summarize(predicted = weighted.mean(predicted, n))

fig_recr <- ggplot(recr_pred_pop, aes(age, predicted)) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), 
              alpha = 0.2) +
  geom_line(linewidth = 1.2, color = "grey50") +
  geom_line(data = recr_pred_wgt, 
            color = "cornflowerblue", 
            linewidth = 1.2) +
  geom_pointrange(aes(y = perc_recr, 
                      ymin = perc_recr - se_recr, 
                      ymax = perc_recr + se_recr),
                  surv_recr_summ) +
  scale_y_continuous("Offspring recruitment", 
                     labels = scales::percent,
                     limits = c(0, 0.62)) +
  labs(x = "Age (years)") +
  theme(legend.position = "none")

@@

sex_ratio_mod <- glmer(
  is_male ~ poly(age10, 2, raw = TRUE) + (1 | animalID) + (1 | year_fct),
  pup_sex_data,
  family = "binomial"
)

summary(sex_ratio_mod)$coefficients

@@

# Create POPULATION-LEVEL (i.e. without rand effects) predictions
sexrat_pred_pop <- ggpredict(
  sex_ratio_mod, 
  terms = "age10 [all]",
  type = "fixed",
  interval = "confidence"
) %>% 
  as_tibble() %>% 
  mutate(age = x * 10) %>% 
  filter(age <= 20)

# Create WEIGHTED-AVERAGE predictions using random effects
seal_years <- count(pup_sex_data, year_fct)
sexrat_pred_wgt <- ggpredict(
  sex_ratio_mod,
  terms = c("age10 [all]", "year_fct [all]"),
  type = "random",
  interval = "confidence"
) %>% 
  as_tibble() %>% 
  mutate(age = x * 10,
         year_fct = group) %>% 
  filter(age <= 20) %>% 
  left_join(seal_years, by = "year_fct") %>% 
  group_by(age) %>% 
  summarize(pred_sd = sd(predicted),
            predicted = weighted.mean(predicted, n))

ggplot(sexrat_pred_pop, aes(age, predicted)) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), 
              alpha = 0.2) +
  geom_line(linewidth = 1.2, color = "grey50") +
  geom_pointrange(aes(y = perc_male, 
                      ymin = perc_male - se, 
                      ymax = perc_male + se),
                  sexrat_summ) +
  scale_y_continuous("Male offspring", labels = scales::percent) +
  labs(x = "Age (years)") +
  theme(legend.position = "none")
```

@@

### Supplemental Material

```{r}
#| label: seg-breeding-mod

# Fit GLMMs to the data with binary variable indicating before and after onset # of senescence. We a priori chose age 11 for senescence (age_senesce).
breed_age_seg <- glmer(
  observed_int ~ age10 * age_cat + (1 | animalID) + (1 | year_fct),
  sealdat,
  family = "binomial",
  control = glmerControl(optimizer = "bobyqa")
)

summary(breed_age_seg)
```

```{r}
#| label: fig-seg-breeding-mod

breed_pred_seg <- ggpredict(breed_age_seg, 
                            terms = c("age10 [all]", "age_cat")) %>% 
  as_tibble() %>% 
  mutate(age = x * 10,
         age_cat = factor(group, levels = c("Young", "Old"))) %>% 
  filter((age_cat == "Young" & age < age_senesce) |
           (age_cat == "Old" & age >= age_senesce))

ggplot(breed_pred_seg, aes(age, predicted)) +
  geom_ribbon(aes(fill = age_cat, ymin = conf.low, ymax = conf.high), 
              alpha = 0.2) +
  geom_line(aes(color = age_cat), linewidth = 1.2) +
  geom_pointrange(aes(y = perc_breed, 
                      ymin = perc_breed - se, 
                      ymax = perc_breed + se),
                  breed_nonbreed) +
  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
  scale_color_manual(values = c("cornflowerblue", "firebrick")) +
  scale_fill_manual(values = c("cornflowerblue", "firebrick")) +
  scale_y_continuous("Breeding", labels = scales::percent) +
  labs(x = "Age (years)") +
  theme(legend.position = "none")

@@

```{r}
#| label: seg-repro-mod

# Fit pup *survival* model
pup_surv_seg <- glmer(
  pup_surv_int ~ age10 * age_cat + (1 | animalID) + (1 | year),
  surv_recr_data,
  family = "binomial",
  control = glmerControl(optimizer = "bobyqa")
)

# Fit pup *recruitment* model
pup_recr_seg <- glmer(
  pup_recr_int ~ age10 * age_cat + (1 | animalID) + (1 | year),
  filter(surv_recr_data, pupsex == "F"),
  family = "binomial",
  control = glmerControl(optimizer = "bobyqa")
)

"Survival"
summary(pup_surv_seg)$coefficients
"Recruitment"
summary(pup_recr_seg)$coefficients

@@

```{r}
#| label: fig-seg-repro
#| fig-width: 9

surv_pred_seg <- ggpredict(pup_surv_seg, 
                           terms = c("age10 [all]", "age_cat")) %>% 
  as_tibble() %>% 
  mutate(age = x * 10,
         age_cat = factor(group, levels = c("Young", "Old"))) %>% 
  filter((age_cat == "Young" & age < age_senesce) |
           (age_cat == "Old" & age >= age_senesce))

fig_surv <- ggplot(surv_pred_seg, aes(age, predicted)) +
  geom_ribbon(aes(fill = age_cat, ymin = conf.low, ymax = conf.high), 
              alpha = 0.2) +
  geom_line(aes(color = age_cat), linewidth = 1.2) +
  geom_pointrange(aes(y = perc_surv, 
                      ymin = perc_surv - se_surv, 
                      ymax = perc_surv + se_surv),
                  surv_recr_summ) +
  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
  scale_color_manual(values = c("cornflowerblue", "firebrick")) +
  scale_fill_manual(values = c("cornflowerblue", "firebrick")) +
  scale_y_continuous("Offspring survival", labels = scales::percent) +
  labs(x = "Age (years)") +
  theme(legend.position = "none")

recr_pred_seg <- ggpredict(pup_recr_seg, 
                           terms = c("age10 [all]", "age_cat")) %>% 
  as_tibble() %>% 
  mutate(age = x * 10,
         age_cat = factor(group, levels = c("Young", "Old"))) %>% 
  filter((age_cat == "Young" & age < age_senesce) |
           (age_cat == "Old" & age >= age_senesce))

fig_recr <- ggplot(recr_pred_seg, aes(age, predicted)) +
  geom_ribbon(aes(fill = age_cat, ymin = conf.low, ymax = conf.high), 
              alpha = 0.2) +
  geom_line(aes(color = age_cat), linewidth = 1.2) +
  geom_pointrange(aes(y = perc_recr, 
                      ymin = perc_recr - se_recr, 
                      ymax = perc_recr + se_recr),
                  surv_recr_summ) +
  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
  scale_color_manual(values = c("cornflowerblue", "firebrick")) +
  scale_fill_manual(values = c("cornflowerblue", "firebrick")) +
  scale_y_continuous("Offspring recruitment", labels = scales::percent) +
  labs(x = "Age (years)") +
  theme(legend.position = "none")

@@

```{r}
#| label: seg-sexrat-mod

# Fit GLMMs to the data with binary variable indicating before and after onset 
# of senescence. We a priori chose age 11 for senescence (age_senesce).
sexrat_age_seg <- glmer(
  is_male ~ age10 * age_cat + (1 | animalID) + (1 | year_fct),
  pup_sex_data,
  family = "binomial",
  control = glmerControl(optimizer = "bobyqa")
)

summary(sexrat_age_seg)$coefficients
```

```{r}
#| label: fig-seg-sexrat

sexrat_pred_seg <- ggpredict(sexrat_age_seg, 
                             terms = c("age10 [all]", "age_cat")) %>% 
  as_tibble() %>% 
  mutate(age = x * 10,
         age_cat = factor(group, levels = c("Young", "Old"))) %>% 
  filter((age_cat == "Young" & age < age_senesce) |
           (age_cat == "Old" & age >= age_senesce))

ggplot(sexrat_pred_seg, aes(age, predicted)) +
  geom_ribbon(aes(fill = age_cat, ymin = conf.low, ymax = conf.high), 
              alpha = 0.2) +
  geom_line(aes(color = age_cat), linewidth = 1.2) +
  geom_pointrange(aes(y = perc_male, 
                      ymin = perc_male - se, 
                      ymax = perc_male + se),
                  sexrat_summ) +
  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
  scale_color_manual(values = c("cornflowerblue", "firebrick")) +
  scale_fill_manual(values = c("cornflowerblue", "firebrick")) +
  scale_y_continuous("Breeding", labels = scales::percent) +
  labs(x = "Age (years)") +
  theme(legend.position = "none")

Supplemental Table 1: Statistics for breeding success model.

```{r, code = FALSE}
#| label: tbl-breed-cap

tbl_breed_cap <- str_glue("The results from breeding probability analysis. The random effect variance of animal id is {format_signif(VarCorr(breeding_mod)$animalID, 3)} and of year was {format_signif(VarCorr(breeding_mod)$year, 3)}.")

```{r, code = FALSE}
#| label: tbl-breed-success
tbl-cap: !expr "tbl_breed_cap"

breed_mod_summary$coefficients %>% 
  as_tibble() %>% 
  mutate(Parameter = c("Intercept", "Age (linear)", "Age (quadratic)"),
         across(c(Estimate, `Std. Error`, `z value`), 
                \(x) format_signif(x, digits = 3)),
         `p value` = format_pval(`Pr(>|z|)`, digits = 2)) %>%
  select(-`Pr(>|z|)`) %>% 
  relocate(Parameter) %>% 
  knitr::kable()
```

```{r, code = FALSE}
#| label: tbl-survrecruit-cap

surv_recruit_cap <- str_glue("The results from offspring survival and recruitment analyses. In the offspring survival model, the random effect variance of animal id is {format_signif(VarCorr(pup_surv_model)$animalID, 3)} and of year was {format_signif(VarCorr(pup_surv_model)$year, 3)}. In the offspring recruitment model, the random effect variance of animal id is {format_signif(VarCorr(pup_recruit_model)$animalID, 3)} and of year was {format_signif(VarCorr(pup_recruit_model)$year, 3)}")

Supplemental table for offspring survival

```{r, code = FALSE}
#| label: tbl-surv_recruit
tbl-cap: !expr "surv_recruit_cap"

surv_mod_summary <- summary(pup_surv_model)
surv_mod_kable <- surv_mod_summary$coefficients %>% 
  as_tibble() %>% 
  mutate(Parameter = c("Intercept", "Age (linear)", "Age (quadratic)"),
         across(c(Estimate, `Std. Error`, `z value`), 
                \(x) format_signif(x, digits = 3)),
         `p value` = format_pval(`Pr(>|z|)`, digits = 2),
         Response = "Survival") %>%
  select(-`Pr(>|z|)`) %>% 
  relocate(Response, Parameter)

recruit_mod_summary <- summary(pup_recruit_model)
recruit_mod_kable <- recruit_mod_summary$coefficients %>% 
  as_tibble() %>% 
  mutate(Parameter = c("Intercept", "Age (linear)", "Age (quadratic)"),
         across(c(Estimate, `Std. Error`, `z value`), 
                \(x) format_signif(x, digits = 3)),
         `p value` = format_pval(`Pr(>|z|)`, digits = 2),
         Response = "Recruitment") %>%
  select(-`Pr(>|z|)`) %>% 
  relocate(Response, Parameter)

rbind(surv_mod_kable, recruit_mod_kable) %>% 
  knitr::kable() %>% 
  kableExtra::kable_styling() %>% 
  kableExtra::collapse_rows(columns = 1)

Supplemental table for pup sex ratio model.

```{r, code = FALSE}
#| label: tbl-pupsex-cap

pupsex_cap <- str_glue("The results from offspring sex ratio analyses. In the model, the random effect variance of animal id is {format_signif(VarCorr(pup_sex_mod)$animalID, 3)} and of year was {format_signif(VarCorr(pup_sex_mod)$year, 3)}.") 
```

```{r, code = FALSE}
#| label: tbl-pupsex
tbl-cap: !expr "pupsex_cap"

pupsex_mod_summary <- summary(pup_sex_mod)
pupsex_quad_summary <- summary(pup_sex_quad)
pupsex_mod_summary$coefficients %>% 
  as_tibble() %>% 
  mutate(Parameter = c("Intercept", "Age (linear)"),
         across(c(Estimate, `Std. Error`, `z value`), 
                \(x) format_signif(x, digits = 3)),
         `p value` = format_pval(`Pr(>|z|)`/2, digits = 3)) %>%
  select(-`Pr(>|z|)`) %>% 
  relocate(Parameter) %>% 
  knitr::kable()

# We fit a quadratic equation, but only the linear term was significant, so the linear model is the one that is plotted. take the p = 0.517 and divide by 2 because a priori. report in supplemental table that its reporting on 1 tailed (no absolute value of z, just z)
```

Supplemental table for male biased sex ratios at older ages.

```{r, code = FALSE}
#| label: tbl-pup-sex
#| tbl-cap: "Offspring sex ratio skewed towards males with increasing maternal age. Mothers age 12, 13, and 17 gave birth to significantly more male than female pups (bold)."

pup_sex_tbl <- pup_sex_summary %>% 
  transmute(
    Age = age,
    `Male pups` = M,
    `Total pups` = num,
    `% male` = scales::label_percent(accuracy = 0.1)(percm),
    `p value` = format_pval(pval, 3)
  )

is_signif <- which(pup_sex_tbl$`p value` < "0.05")

knitr::kable(pup_sex_tbl) %>% 
  kableExtra::kable_styling() %>% 
  kableExtra::row_spec(is_signif, bold = TRUE)
```

Supplemental figure for phenology models

```{r, code = FALSE}
#| fig-width: 8

#NOTE: Both BA and BD have some points that are really late (above 300 doy) that need fixing. Maybe make boxplots also..?
seal_phenology %>% 
  rename(`Breeding\narrival` = BA,
         `Breeding\ndeparture` = BD,
         `Molt\narrival` = MA,
         `Molt\ndeparture` = MD) %>% 
  pivot_longer(`Breeding\narrival`:`Molt\ndeparture`, 
               names_to = "event", 
               values_to = "DOY") %>% 
ggplot(aes(y = DOY)) +
  geom_boxplot(aes(factor(age))) +
  geom_smooth(aes(as.integer(factor(age))), method = "lm", formula = y ~ x, se = TRUE) +
  facet_wrap(~ event, nrow = 2, scales = "free_y") +
  labs(x = "Age (years)", y = "Day of year") +
  theme_classic()
```

Supplemental table for the phenology models.

```{r, code = FALSE}
ba_data <- phenology_models$model_data[[1]]
ba_mod <- lmer(doy ~ age + (1 | animalID) + (1 | year), data = ba_data)
ba_summary <- summary(ba_mod)
ba_ktable <- ba_summary$coefficients %>% 
  as_tibble() %>% 
  mutate(Parameter = c("Intercept", "Age (linear)"),
         across(c(Estimate, `Std. Error`, `df`, `t value`), 
                \(x) format_signif(x, digits = 3)),
         `p value` = format_pval(`Pr(>|t|)`, digits = 2),
         Response = "Breeding arrival") %>%
  select(-`Pr(>|t|)`) %>% 
  relocate(Response, Parameter)

bd_data <- phenology_models$model_data[[2]]
bd_mod <- lmer(doy ~ age + (1 | animalID) + (1 | year), data = ba_data)
bd_summary <- summary(ba_mod)
bd_ktable <- bd_summary$coefficients %>% 
  as_tibble() %>% 
  mutate(Parameter = c("Intercept", "Age (linear)"),
         across(c(Estimate, `Std. Error`, `df`, `t value`), 
                \(x) format_signif(x, digits = 3)),
         `p value` = format_pval(`Pr(>|t|)`, digits = 2),
         Response = "Breeding departure") %>%
  select(-`Pr(>|t|)`) %>% 
  relocate(Response, Parameter)

ma_data <- phenology_models$model_data[[3]]
ma_mod <- lmer(doy ~ age + (1 | animalID) + (1 | year), data = ma_data)
ma_summary <- summary(ma_mod)
ma_ktable <- ma_summary$coefficients %>% 
  as_tibble() %>% 
  mutate(Parameter = c("Intercept", "Age (linear)"),
         across(c(Estimate, `Std. Error`, `df`, `t value`), 
                \(x) format_signif(x, digits = 3)),
         `p value` = format_pval(`Pr(>|t|)`, digits = 2),
         Response = "Molt arrival") %>%
  select(-`Pr(>|t|)`) %>% 
  relocate(Response, Parameter)

md_data <- phenology_models$model_data[[4]]
md_mod <- lmer(doy ~ age + (1 | animalID) + (1 | year), data = md_data)
md_summary <- summary(md_mod)
md_ktable <- md_summary$coefficients %>% 
  as_tibble() %>% 
  mutate(Parameter = c("Intercept", "Age (linear)"),
         across(c(Estimate, `Std. Error`, `df`, `t value`), 
                \(x) format_signif(x, digits = 3)),
         `p value` = format_pval(`Pr(>|t|)`, digits = 2),
         Response = "Molt departure") %>%
  select(-`Pr(>|t|)`) %>% 
  relocate(Response, Parameter)

rbind(ba_ktable, bd_ktable, ma_ktable, md_ktable) %>% 
  knitr::kable() %>% 
  kableExtra::kable_styling() %>% 
  kableExtra::collapse_rows(columns = 1)
```

>>>>

### H1: Reproductive senescence

@@

# Fit GLMMs to the data with binary variable indicating before and after onset # of senescence. We a priori chose age 11 for senescence (age_senesce).
breed_age_mod <- glmer(
  observed_int ~ age10 * age_cat + (1 | animalID) + (1 | year_fct),
  sealdat,
  family = "binomial",
  control = glmerControl(optimizer = "bobyqa")
)

@@

```{r}
#| label: fig-breeding-mod
#| fig-cap: "Breeding rates increased until age 10, then decreased from age 11 on. Black points represent mean and standard error of the observed breeding rates. The solid lines and shaded areas represent the mean and 95% confidence interval of the population-level model predictions. The observed data are unequally distributed across years, so the dashed lines indicate the predicted breeding rates averaged over years (i.e., including the random effect of year), weighted by the number of seals observed in each year."

@@

# Create POPULATION-LEVEL (i.e. without rand effects) predictions
breed_pred_pop <- ggpredict(breed_age_mod, 
                            terms = c("age10 [all]", "age_cat")) %>% 
  as_tibble() %>% 
  mutate(age = x * 10,
         age_cat = factor(group, levels = c("Young", "Old"))) %>% 
  filter((age_cat == "Young" & age < age_senesce) |
           (age_cat == "Old" & age >= age_senesce),
         age <= 20)

# Create WEIGHTED-AVERAGE predictions using years' random effects
seal_years <- count(sealdat, year_fct)
breed_pred_wgt <- ggpredict(
  breed_age_mod,
  terms = c("age10 [all]", "age_cat", "year_fct [all]"),
  type = "random",
  interval = "confidence"
) %>% 
  as_tibble() %>% 
  mutate(age = x * 10,
         age_cat = group,
         year_fct = facet) %>% 
  filter((age_cat == "Young" & age < age_senesce) |
           (age_cat == "Old" & age >= age_senesce),
         age <= 20) %>% 
  left_join(seal_years, by = "year_fct") %>% 
  group_by(age, age_cat) %>% 
  summarize(predicted = weighted.mean(predicted, n),
            .groups = "drop")

ggplot(breed_pred_pop, aes(age, predicted)) +
  geom_ribbon(aes(fill = age_cat, ymin = conf.low, ymax = conf.high), 
              alpha = 0.2) +
  geom_line(aes(color = age_cat), linewidth = 1.2) +
  geom_line(aes(color = age_cat),
            breed_pred_wgt, 
            linewidth = 1.2,
            linetype = "dashed") +
  geom_pointrange(aes(y = perc_breed, 
                      ymin = perc_breed - se, 
                      ymax = perc_breed + se),
                  breed_nonbreed) +
  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
  scale_y_continuous("Breeding", labels = scales::percent) +
  scale_color_manual(values = c("cornflowerblue", "firebrick")) +
  scale_fill_manual(values = c("cornflowerblue", "firebrick")) +
  labs(x = "Age (years)") +
  theme(legend.position = "none")

@@

# Isolate pup survival/recruitment 
surv_recr_data <- sealdat %>% 
  drop_na(pupID) %>% 
  # Only pups born 2016 or earlier (otherwise survival biased)
  filter(pupyearborn <= 2016) %>% 
  mutate(pup_surv_int = as.integer(pup_survived),
         pup_recr_int = as.integer(puprecruited))

# Fit pup *survival* model
pup_surv_mod <- glmer(
  pup_surv_int ~ age10 * age_cat + (1 | animalID) + (1 | year_fct),
  surv_recr_data,
  family = "binomial",
  control = glmerControl(optimizer = "bobyqa")
)

# Fit pup *recruitment* model (female offspring only)
pup_recr_mod <- glmer(
  # NOTE: no random effect for individual bc singular fit
  # In survival model, 93 individuals (15%) had >3 pups (male+female pups)
  # In recruit model, only 17 individuals (4%) had >3 pups (female pups only)
  pup_recr_int ~ age10 * age_cat + (1 | year_fct),
  filter(surv_recr_data, pupsex == "F"),
  family = "binomial",
  control = glmerControl(optimizer = "bobyqa")
)

@@

# Create pup survival POPULATION-LEVEL (i.e. without rand effects) predictions
surv_pred_pop <- ggpredict(
  pup_surv_mod, 
  terms = c("age10 [all]", "age_cat"),
  type = "fixed",
  interval = "confidence"
) %>% 
  as_tibble() %>% 
  mutate(age = x * 10,
         age_cat = group) %>% 
  filter((age_cat == "Young" & age < age_senesce) |
           (age_cat == "Old" & age >= age_senesce),
         age <= 20)

fig_surv <- ggplot(surv_pred_pop, aes(age, predicted)) +
  geom_ribbon(aes(fill = age_cat, ymin = conf.low, ymax = conf.high), 
              alpha = 0.2) +
  geom_line(aes(color = age_cat), linewidth = 1.2) +
  geom_pointrange(aes(y = perc_surv, 
                      ymin = perc_surv - se_surv, 
                      ymax = perc_surv + se_surv),
                  surv_recr_summ) +
  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
  scale_y_continuous("Offspring survival", 
                     labels = scales::percent,
                     limits = c(0, 0.62)) +
  scale_color_manual(values = c("cornflowerblue", "firebrick")) +
  scale_fill_manual(values = c("cornflowerblue", "firebrick")) +
  labs(x = "Age (years)") +
  theme(legend.position = "none")

# Create pup recruit POPULATION-LEVEL (i.e. without rand effects) predictions
recr_pred_pop <- ggpredict(
  pup_recr_mod, 
  terms = c("age10 [all]", "age_cat"),
  type = "fixed",
  interval = "confidence"
) %>% 
  as_tibble() %>% 
  mutate(age = x * 10,
         age_cat = group) %>% 
  filter((age_cat == "Young" & age < age_senesce) |
           (age_cat == "Old" & age >= age_senesce),
         age <= 20)

fig_recr <- ggplot(recr_pred_pop, aes(age, predicted)) +
  geom_ribbon(aes(fill = age_cat, ymin = conf.low, ymax = conf.high), 
              alpha = 0.2) +
  geom_line(aes(color = age_cat), linewidth = 1.2) +
  geom_pointrange(aes(y = perc_recr, 
                      ymin = perc_recr - se_recr, 
                      ymax = perc_recr + se_recr),
                  surv_recr_summ) +
  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
  scale_y_continuous("Offspring recruitment", 
                     labels = scales::percent,
                     limits = c(0, 0.62)) +
  scale_color_manual(values = c("cornflowerblue", "firebrick")) +
  scale_fill_manual(values = c("cornflowerblue", "firebrick")) +
  labs(x = "Age (years)") +
  theme(legend.position = "none")

@@

sexrat_age_mod <- glmer(
  is_male ~ age10 * age_cat + (1 | animalID) + (1 | year_fct),
  pup_sex_data,
  family = "binomial"
)

summary(sexrat_age_mod)$coefficients

@@

# Create POPULATION-LEVEL (i.e. without rand effects) predictions
sexrat_pred <- ggpredict(sexrat_age_mod, 
                         terms = c("age10 [all]", "age_cat")) %>% 
  as_tibble() %>% 
  mutate(age = x * 10,
         age_cat = factor(group, levels = c("Young", "Old"))) %>% 
  filter((age_cat == "Young" & age < age_senesce) |
           (age_cat == "Old" & age >= age_senesce))

ggplot(sexrat_pred, aes(age, predicted)) +
  geom_ribbon(aes(fill = age_cat, ymin = conf.low, ymax = conf.high), 
              alpha = 0.2) +
  geom_line(aes(color = age_cat), linewidth = 1.2) +
  geom_pointrange(aes(y = perc_male, 
                      ymin = perc_male - se, 
                      ymax = perc_male + se),
                  sexrat_summ) +
  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
  scale_color_manual(values = c("cornflowerblue", "firebrick")) +
  scale_fill_manual(values = c("cornflowerblue", "firebrick")) +
  scale_y_continuous("Male offspring", labels = scales::percent) +
  labs(x = "Age (years)") +
  theme(legend.position = "none")
```

@@

## Supplemental Material

### Models' sensitivity to age of senescence

```{r}
#| label: mod-sens

fit_models <- function(a) {
  age_senesce <- a
  sealdat <- sealdat %>% 
    mutate(age_cat = factor(age >= age_senesce, 
                            labels = c("Young", "Old")))

  # H1
  breed_age_mod <- glmer(
    observed_int ~ age10 * age_cat + (1 | animalID) + (1 | year_fct),
    sealdat,
    family = "binomial",
    control = glmerControl(optimizer = "bobyqa")
  )

  # H2
  surv_recr_data <- sealdat %>% 
    drop_na(pupID) %>% 
    filter(pupyearborn <= 2016) %>% 
    mutate(pup_surv_int = as.integer(pup_survived),
           pup_recr_int = as.integer(puprecruited))

  pup_surv_mod <- glmer(
    pup_surv_int ~ age10 * age_cat + (1 | animalID) + (1 | year_fct),
    surv_recr_data,
    family = "binomial",
    control = glmerControl(optimizer = "bobyqa")
  )

  pup_recr_mod <- glmer(
    pup_recr_int ~ age10 * age_cat + (1 | year_fct),
    filter(surv_recr_data, pupsex == "F"),
    family = "binomial",
    control = glmerControl(optimizer = "bobyqa")
  )

  # H3
  pup_sex_data <- sealdat %>% 
    filter(pupsex %in% c("M", "F"), 
           age < 19) %>% 
    mutate(is_male = as.integer(pupsex == "M"))

  sexrat_age_mod <- glmer(
    is_male ~ age10 * age_cat + (1 | animalID) + (1 | year_fct),
    pup_sex_data,
    family = "binomial"
  )

  list(
    h1 = summary(breed_age_mod)$coefficients,
    h2_surv = summary(pup_surv_mod)$coefficients,
    h2_recr = summary(pup_recr_mod)$coefficients,
    h3 = summary(sexrat_age_mod)$coefficients
  )
}

# Fit models with cut-off at ages 8-13
cutoff <- 8:13
mod_sens <- map(cutoff, fit_models)

# Are hypotheses supported by the model?
extract_p <- \(h) map_dbl(mod_sens, \(m) m[[h]]["age10:age_catOld", "Pr(>|z|)"])
p_tbl <- tibble(age_cutoff = cutoff, 
                "H1 (breeding)" = extract_p("h1"),
                "H2a (survival)" = extract_p("h2_surv"),
                "H2b (recruitment)" = extract_p("h2_recr"),
                "H3 (sex ratio)" = extract_p("h3"))

knitr::kable(p_tbl)

@@

```{r}
#| label: fig-mod-sens
#| fig-cap: "Model sensitivity to onset of senescence. Age at onset of senescence on the x-axis, p-value for the coefficient of the interaction (age category = post-senescence) x (response variable) on the y-axis. Statistical significance (p = 0.05) indicated by dashed line. Hypotheses H1 (reproductive senescence), H2b (maternal-effect senescence; offspring recruitment), and H3 (offspring sex ratio) were insensitive to choice of age for onset of senescence. Hypothesis H2a (maternal-effect senescence; offspring survival) was significantly supported for onset of senescence age 11+. At age 10, the p-vale for the coefficient of the interaction (age category = post-senescence) x (offspring survival) was 0.051."

p_tbl %>% 
  pivot_longer(-age_cutoff, names_to = "hypothesis", values_to = "p") %>% 
  ggplot(aes(age_cutoff, p)) +
  geom_point() +
  geom_hline(yintercept = 0.05, linetype = "dashed") +
  scale_y_log10(labels = scales::number) +
  labs(x = "Senescence onset (years)") +
  facet_wrap(~hypothesis)

@@

### Haul-out duration by reproductive state

```{r}
#| label: fig-haul-dur

haul_dur <- drop_na(sealdat, breeddur)
haul_summ <- haul_dur %>% 
  group_by(observed) %>% 
  summarize(breeddur_10 = quantile(breeddur, 0.1),
            breeddur_lbl = str_glue("10% < {breeddur_10} days"))
ggplot(haul_dur, aes(breeddur)) +
  geom_density(aes(fill = observed), alpha = 0.5) +
  geom_density(aes(color = observed)) +
  geom_vline(aes(color = observed, xintercept = breeddur_10), 
             haul_summ) +
  geom_text(aes(x = breeddur_10 + 0.5, label = breeddur_lbl), 
            haul_summ, 
            y = 0.09,
            hjust = 0) +
  scale_x_continuous("Haul-out duration (days)", 
                     limits = c(0, 50)) +
  scale_color_manual(values = c("darkgreen", "darkorchid")) +
  scale_fill_manual(values = c("darkgreen", "darkorchid")) +
  theme(legend.position = "none")

@@
